---
title: "Results"
author: "Philippe Berends, 3972003"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: true
---

This script contains the code needed to reproduce the plots and tables in my
thesis. In order to run this code, you require the .rds files generated from
the MC experiments.
:

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.path = "plots/",
  dev = "png",
  dpi = 300,
  fig.width = 6,
  fig.height = 4,
  fig.retina = 2
)
```

My experiment was called "ResultRun", so I will use this name to gather files.
If you were to reproduce these results, you should change the string in this
block to the one used in generating the results

```{r}
exp_name = "FinalRun"
```


# Load libraries

```{r}
library(tidyverse)
library(survival)
library(glue)
library(xtable)
library(ggpattern)
library(patchwork)
library(survminer)
```

# Thesis generic

## $\rho$ values

```{r}
frho = function(rho, rho_between = 0, p_g){
  # Function used to Rho correlations
  return(pmax(rho^((0:(p_g-1))/(p_g-1)), rep(rho_between, p_g)))
}

# Various examples
small = frho(rho = 0.2, 0, 100)
medium = frho(rho = 0.5, 0, 100)
large = frho(rho = 0.8, 0, 100)
xl = frho(rho = 0.9, 0, 100)

large_b = frho(rho = 0.4, 0.6, 100)

plotframe = data.frame(small,medium,large,xl, large_b)




# Create long format for ggplot
plotframe = stack(plotframe)

plotframe$ind = factor(plotframe$ind, 
                      levels = c("small", "medium", "large", "xl", "large_b"),
                      labels = c("ρw = 0.2, ρb = 0", 
                                 "ρw = 0.5, ρb = 0",
                                 "ρw = 0.8, ρb = 0", 
                                 "ρw = 0.9, ρb = 0",
                                 "ρw = 0.4, ρb = 0.6"))


# Index column
plotframe["index"] = rep(1:100, 5)

colors = RColorBrewer::brewer.pal(5, "Set1")

plotframe %>%
  ggplot(aes(y = values, group = ind, x = index, colour = ind)) +
  geom_line() +
  labs(title = "Value of f(ρw, ρb) for a group of p/g = 100",
     x = "Index", 
     y = "ρ",
     color = "ρ-values") +
  scale_color_brewer(palette = "Set1") + 
  theme_classic() +
  # Intercepts for the rho values
  geom_hline(yintercept = 0.2, linewidth = 0.4, color = colors[1], linetype = "dashed") +
  geom_hline(yintercept = 0.5, linewidth = 0.4, color = colors[2], linetype = "dashed") +
  geom_hline(yintercept = 0.8, linewidth = 0.4, color = colors[3], linetype = "dashed") +
  geom_hline(yintercept = 0.9, linewidth = 0.4, color = colors[4], linetype = "dashed") +
  geom_hline(yintercept = 0.4, linewidth = 0.4, color = colors[5], linetype = "dashed") +
  geom_hline(yintercept = 0.6, linewidth = 0.4, color = colors[5], linetype = "dashed")
```

# Cauchy shape given p

(In progress)

```{r}
qcauchy(seq(-1,1,0.001), scale = 0.1) %>% plot()
```


# MC runs

## Import and parse results

```{r}
# Fetch files containing exp_name
experiment_files = list.files(pattern = glue("{exp_name}.*.rds"))

# Fetch csv with parameter grid
experiment_grid = list.files(pattern = glue("Grid.*{exp_name}.csv"))
experiment_grid = read.csv(experiment_grid, header = T)
```


```{r}
results_methods = data.frame()

for (i in 1:nrow(experiment_grid)) {
  exp_I = glue("Experiment:_FinalRun_giter_{i}.rds")
  
  
  experimentsMC = read_rds(exp_I)
  M = length(experimentsMC$mc_results)
  
  
  gridexp_bayes_rmses = data.frame()
  gridexp_freq_rmses = data.frame()
  gridexp_time = data.frame()
  for (m in 1:M) {
    timing = experimentsMC$mc_results[[m]]$timing
    
    # Extract RMSEs
    brms_rmses = experimentsMC$mc_results[[m]]$results$brms$rmses %>%
      as.data.frame()
    freq_rmses = experimentsMC$mc_results[[m]]$results$aft$rmses %>%
      as.data.frame()
    
    # If either model failed to fit (NA)
    if (mean(dim(freq_rmses) == c(0,0)) == 1) {
      freq_rmses = data.frame(coef = NA, mrl = NA, bioage = NA)
    }
    if (mean(dim(brms_rmses) == c(0,0)) == 1) {
      brms_rmses = data.frame(coef = NA, mrl = NA, bioage = NA)
    }
    gridexp_bayes_rmses = rbind(gridexp_bayes_rmses, brms_rmses)
    gridexp_freq_rmses = rbind(gridexp_freq_rmses, freq_rmses)
    gridexp_time = rbind(gridexp_time, timing) # Bayesian = 2 freq = 1
  }
  
  # gridexp_nerr_bayes = gridexp_bayes_rmses[,1] %>% is.na() %>% sum()
  # gridexp_nerr_freq = gridexp_freq_rmses[,1] %>% is.na() %>% sum()
  
  # Record performance
  gridexp_param = experimentsMC$params
  colnames(gridexp_param)[5] = "method"
  # Add frequentist as a factor level for method
  levels(gridexp_param$method) = c("horseshoe", "cauchy", "frequentist")
  
  gridexp_param = rbind(gridexp_param, gridexp_param)
  # Second row is frequentist - first bayesian
  gridexp_param[2,]$method = as.factor("frequentist")
  
  ## Add scores
  # Success%
  gridexp_param[1,"psuccess"] = (!gridexp_bayes_rmses[,1] %>% is.na()) %>% mean()
  gridexp_param[2,"psuccess"] = (!gridexp_freq_rmses[,1] %>% is.na()) %>% mean()
  
  # Means
  gridexp_param[1,c("mu_RMSE_coef","mu_RMSE_mrl", "mu_RMSE_bioage")] = colMeans(gridexp_bayes_rmses,na.rm = T)
  gridexp_param[2,c("mu_RMSE_coef","mu_RMSE_mrl", "mu_RMSE_bioage")] = colMeans(gridexp_freq_rmses,na.rm = T)
  
  # Variances
  gridexp_param[1,c("var_RMSE_coef","var_RMSE_mrl", "var_RMSE_bioage")] = 
    apply(gridexp_bayes_rmses, 2, var, na.rm = T)
  
  gridexp_param[2,c("var_RMSE_coef","var_RMSE_mrl", "var_RMSE_bioage")] = 
    apply(gridexp_freq_rmses, 2, var, na.rm = T)
  
  # Timing, tricky due to errors; time is of non-failed runs!
  # Each run has a time, whether failed or not
  bsuccesses = !gridexp_bayes_rmses[,1] %>% is.na()
  fsuccesses = !gridexp_freq_rmses[,1] %>% is.na()
  
  gridexp_param[1,"mu_success_time_sec"] = mean(gridexp_time[bsuccesses,2])
  gridexp_param[2,"mu_success_time_sec"] = mean(gridexp_time[fsuccesses,1])
  
  gridexp_param[1,"var_success_time_sec"] = var(gridexp_time[bsuccesses,2])
  gridexp_param[2,"var_success_time_sec"] = var(gridexp_time[fsuccesses,1])

  results_methods = rbind(results_methods, gridexp_param)
}
# Remove rownames added by rbind
rownames(results_methods) = NULL

# keep unique frequentist experiments (were performed twice)
results_methods = results_methods %>%
    distinct(n_obs, p, g, non_zero, method, .keep_all = TRUE)
```

## RMSE plots



```{r}
results_methods %>%
  filter(non_zero == 1) %>%
  filter(g == 4) %>%
  # To 1 col
  pivot_longer(cols = c(mu_RMSE_bioage, mu_RMSE_mrl),
               names_to = "type", values_to ="RMSE") %>%
  # Rename
  mutate(
    type = ifelse(type == "mu_RMSE_bioage", "Biological Age", "Mean residual life")
  ) %>%
  # Add variances
  mutate(varcs = ifelse(
    type == "Biological Age", var_RMSE_bioage, var_RMSE_mrl)
  ) %>%
  # plot
  ggplot(aes(x = factor(n_obs), y = RMSE, 
             colour = factor(p), group = factor(p))) +
  geom_point(position = position_dodge(width = 0.5), size = 1) +
  geom_errorbar(aes(ymin = RMSE - sqrt(varcs)/sqrt(psuccess*100), # 100 because M = 100 
                    ymax = RMSE + sqrt(varcs)/sqrt(psuccess*100)),
                position = position_dodge(width = 0.5), 
                width = 0.2) +
  # facet_wrap(~ method) +
  facet_grid(type ~ method, scales = "free_y") +
  labs(title = "Performance of AccelerAge using three methods: Non-sparse case",
     subtitle = expression(g[nz] == 1 ~ " " ~ g == 4),
     x = expression(n[obs]), 
     colour = "p", 
     y = expression()) +
  theme_linedraw() + 
  theme(panel.grid.major = element_blank())
```

```{r}
results_methods %>%
  filter(non_zero == 0.25) %>%
  filter(g == 4) %>%
  # To 1 col
  pivot_longer(cols = c(mu_RMSE_bioage, mu_RMSE_mrl),
               names_to = "type", values_to ="RMSE") %>%
  # Rename
  mutate(
    type = ifelse(type == "mu_RMSE_bioage", "Biological Age", "Mean residual life")
  ) %>%
  mutate(varcs = ifelse(
    type == "Biological Age", var_RMSE_bioage, var_RMSE_mrl)
  ) %>%
  # plot
  ggplot(aes(x = factor(n_obs), y = RMSE, 
             colour = factor(p), group = factor(p))) +
  geom_point(position = position_dodge(width = 0.5), size = 1) +
  geom_errorbar(aes(ymin = RMSE - sqrt(varcs)/sqrt(psuccess*100), # 100 because M = 100 
                    ymax = RMSE + sqrt(varcs)/sqrt(psuccess*100)),
                position = position_dodge(width = 0.5), 
                width = 0.2) +
  # facet_wrap(~ method) +
  facet_grid(type ~ method, scales = "free_y") +
  labs(title = "Performance of AccelerAge using three methods: Sparse case",
     subtitle = expression(g[nz] == 0.25 ~ " " ~ g == 4),
     x = expression(n[obs]), 
     colour = "p", 
     y = expression("RMSE")) +
  theme_linedraw() + 
  theme(panel.grid.major = element_blank())
```

```{r}
results_methods %>%
  filter(non_zero == 1) %>%
  filter(g == 10) %>%
  # To 1 col
  pivot_longer(cols = c(mu_RMSE_bioage, mu_RMSE_mrl),
               names_to = "type", values_to ="RMSE") %>%
  # Rename
  mutate(
    type = ifelse(type == "mu_RMSE_bioage", "Biological Age", "Mean residual life")
  ) %>%
  mutate(varcs = ifelse(
    type == "Biological Age", var_RMSE_bioage, var_RMSE_mrl)
  ) %>%
  # plot
  ggplot(aes(x = factor(n_obs), y = RMSE, 
             colour = factor(p), group = factor(p))) +
  geom_point(position = position_dodge(width = 0.5), size = 1) +
  geom_errorbar(aes(ymin = RMSE - sqrt(varcs)/sqrt(psuccess*100), # 100 because M = 100 
                    ymax = RMSE + sqrt(varcs)/sqrt(psuccess*100)),
                position = position_dodge(width = 0.5), 
                width = 0.2) +
  # facet_wrap(~ method) +
  facet_grid(type ~ method, scales = "free_y") +
  labs(title = "Performance of AccelerAge using three methods: Non-sparse case",
     subtitle = expression(g[nz] == 1 ~ " " ~ g == 10),
     x = expression(n[obs]), 
     colour = "p", 
     y = expression("RMSE")) +
  theme_linedraw() + 
  theme(panel.grid.major = element_blank())
```

```{r}
results_methods %>%
  filter(non_zero == 0.25) %>%
  filter(g == 10) %>%
  # To 1 col
  pivot_longer(cols = c(mu_RMSE_bioage, mu_RMSE_mrl),
               names_to = "type", values_to ="RMSE") %>%
  # Rename
  mutate(
    type = ifelse(type == "mu_RMSE_bioage", "Biological Age", "Mean residual life")
  ) %>%
  mutate(varcs = ifelse(
    type == "Biological Age", var_RMSE_bioage, var_RMSE_mrl)
  ) %>%
  # plot
  ggplot(aes(x = factor(n_obs), y = RMSE, 
             colour = factor(p), group = factor(p))) +
  geom_point(position = position_dodge(width = 0.5), size = 1) +
  geom_errorbar(aes(ymin = RMSE - sqrt(varcs)/sqrt(psuccess*100), # 100 because M = 100 
                    ymax = RMSE + sqrt(varcs)/sqrt(psuccess*100)),
                position = position_dodge(width = 0.5), 
                width = 0.2) +
  # facet_wrap(~ method) +
  facet_grid(type ~ method, scales = "free_y") +
  labs(title = "Performance of AccelerAge using three methods: Sparse case",
     subtitle = expression(g[nz] == 0.25 ~ " " ~ g == 10),
     x = expression(n[obs]), 
     colour = "p", 
     y = expression("RMSE")) +
  theme_linedraw() + 
  theme(panel.grid.major = element_blank())
```


## Auxiliary plots and tables


### Percentiles

```{r}
lp_test = exp(rnorm(1000000, sd = 0.11))
t_test = tibble(linpred = lp_test,
                age_of_death = stats::rweibull(n= length(lp_test),
                                               shape = 15,
                                               scale = 80 * lp_test))

quants = quantile(t_test$linpred, probs = seq(0, 1, 0.2))

t_test = t_test %>%
    mutate(linpred_quantile = cut(
        linpred,
        breaks = quants,
        include.lowest = TRUE,
        labels = paste0(seq(0,80,20), "-", seq(20,100,20),"%")
    ))

t_test %>%
  ggplot(aes(x = age_of_death, group = linpred_quantile, color = linpred_quantile)) +
  geom_density(linewidth = 1.1) +
  theme_linedraw() +
  labs(color = "Percentile\n(Lowest to\nhighest)",
       title = expression("Age-at-death using Weibull according to various percentiles of "
                          * beta^T * italic(X)),
       x = "Chronological age") +
  xlim(c(0, 120))
```


### Fit table


```{r}
library(MASS)
# Remove irrelevant predictors to minimize aic
fit = lm(mu_RMSE_bioage ~ method * p * g * n_obs * non_zero,
         data = results_methods) %>% stepAIC(direction = "both")

xtable(fit %>% summary)
```

### Time

```{r}
results_methods %>% 
  # --- For grouping
  group_by(method, n_obs, p, g, non_zero) %>%
  mutate(p_g = interaction(p, g)) %>%
  # ---
  ggplot(aes(x = as.factor(n_obs), y = mu_success_time_sec,
             fill = as.factor(p),
             pattern = as.factor(g))) +
  geom_col_pattern(aes(group = p_g),
                   position = position_dodge(width = 0.9),
                   pattern_density = 0.01,
                   colour = "black", size = 0.3) +
  geom_errorbar(aes(ymin = mu_success_time_sec - sqrt(var_success_time_sec)/sqrt(psuccess*100),
                    ymax = mu_success_time_sec + sqrt(var_success_time_sec)/sqrt(psuccess*100),
                    group = p_g),
                position = position_dodge(width = 0.9), width = 0.4) +
  scale_pattern_manual(values = c("4" = "none", "10" = "stripe")) +
  scale_fill_brewer(palette = "Set2") +
  facet_grid(non_zero ~ method,
             labeller = labeller(non_zero =
                                   c("0.25" = "g_nz = 0.25",
                                     "1" = "g_nz = 1")
                                 )
             ) +
  labs(x = expression(n[obs]), y = "Mean Fit Time (seconds)",
       fill = "p", pattern = "g") +
  theme_linedraw() +
  theme(legend.position = "bottom",
        panel.grid.major = element_blank())
```


### Success rate

```{r}
results_methods %>% 
  group_by(method, n_obs, p, g, non_zero) %>%
  summarise(psuccess = mean(psuccess)) %>%
  ggplot(aes(x = n_obs, y = psuccess, color = as.factor(p),
             shape = as.factor(g), linetype = as.factor(non_zero),
             group = interaction(as.factor(p), as.factor(g), as.factor(non_zero)))) +
  geom_point(size = 2.5) +
  scale_shape_manual(values=c(6, 18))+
  geom_line() +
  scale_color_brewer(palette = "Dark2") +
  facet_wrap(~method) + 
  labs(title = "Success rate for M = 100 experiments",
      x = expression(n[obs]),
      y = "Success Probability",
      color = "p",
      shape = "g",
      linetype = expression(g[nz])) +
 theme_linedraw()
```


### Shrinkage

Will take the distributions of linpreds for each run

#### Fetch data function

```{r}
tryCatch(
  expr = {
    all_experiments = readRDS("All_experiments.rds")
    all_betas = readRDS("All_betas.rds")
  }, 
  error = function(msg){
    
    # Will store linear predictors
    all_experiments = tibble()
    
    # Will store estimated betas vs actual with parameters
    all_betas = tibble()
    
    for (i in 1:nrow(experiment_grid)) {
      print(glue("Processing experiment {i} out of {nrow(experiment_grid)}"))
      exp_I =  glue("Experiment:_FinalRun_giter_{i}.rds")
      experimentsMC = read_rds(exp_I)
      M = length(experimentsMC$mc_results)
    
      p = experiment_grid[i,]$p
      n = experiment_grid[i,]$n_obs
    
      for (m in 1:M) {
        # idx = ((m-1)*n + 1):(m*n)
        
        brms_res = experimentsMC$mc_results[[m]]$results$brms
        freq_res = experimentsMC$mc_results[[m]]$results$aft
        
        # Check if either Bayesian or frequentist methods managed to converge
        no_error_brms = !is.null(brms_res)
        no_error_freq = !is.null(freq_res)
      
        # Extract the linear predictors for both methods
        if (no_error_brms) {
          X = brms_res$test_data[,1:p]
          
          # Estimated betas
          b_bhats = brms_res$predicted_betas
          # True betas
          b_btrue = brms_res$true_betas$beta
          # Corresponding groups
          b_gtrue = brms_res$true_betas$group
          
          # Get true and estimated linpreds (beta^TX)
          b_h_linpreds = rowSums(sweep(X, 2, b_bhats, "*"))
          true_linpreds = rowSums(sweep(X, 2, b_btrue, "*"))
          
          # Get weibull distribution estimates
          b_shape = brms_res$distr_ests$shape
          b_scale = brms_res$distr_ests$scale
          
          idx = 1:length(b_h_linpreds)
          # Append bayesian rows to all experiment frame
          bayes_row = tibble(
            experiment = i,
            Mc_rep = m,
            id = idx,
            linpreds = b_h_linpreds,
            # true_linpreds = b_t_linpreds,
            method = experiment_grid[i,]$beta_prior,
            n_obs = experiment_grid[i,]$n_obs,
            p = experiment_grid[i,]$p,
            g = experiment_grid[i,]$g,
            non_zero = experiment_grid[i,]$non_zero,
            shape = b_shape,
            scale = b_scale
          )
          all_experiments = bind_rows(all_experiments, bayes_row)
          
          
          idx = 1:length(b_btrue)
          # Append rows to betaframe
          beta_rows = tibble(
            experiment = i,
            Mc_rep = m,
            beta_id = idx,
            true_beta = b_btrue,
            betahat = b_bhats,
            beta_group = b_gtrue,
            method = experiment_grid[i,]$beta_prior,
            n_obs = experiment_grid[i,]$n_obs,
            p = experiment_grid[i,]$p,
            g = experiment_grid[i,]$g,
            non_zero = experiment_grid[i,]$non_zero
          )
          all_betas = bind_rows(all_betas, beta_rows)
          
        }
        
        if (no_error_freq) {
          X = freq_res$test_data[,1:p]
          # Estimated betas
          f_bhats = freq_res$predicted_betas
          # True betas
          f_btrue = freq_res$true_betas$beta
          # Corresponding true groups
          f_gtrue = freq_res$true_betas$group
          
          # Get true and estimated linpreds (beta^TX)
          f_h_linpreds = rowSums(sweep(X, 2, f_bhats, "*"))
          true_linpreds = rowSums(sweep(X, 2, f_btrue, "*"))
          
          # Get weibull distribution estimates
          f_shape = freq_res$distr_ests$shape %>% as.numeric()
          f_scale = freq_res$distr_ests$scale %>% as.numeric()
          
          idx = 1:length(f_h_linpreds)
          # Append frequentist rows to df
          freq_row = tibble(
            experiment = i,
            Mc_rep = m,
            id = idx,
            linpreds = f_h_linpreds,
            # true_linpreds = f_t_linpreds,
            method = "frequentist",
            n_obs = experiment_grid[i,]$n_obs,
            p = experiment_grid[i,]$p,
            g = experiment_grid[i,]$g,
            non_zero = experiment_grid[i,]$non_zero,
            shape = f_shape,
            scale = f_scale
          )
          all_experiments = bind_rows(all_experiments, freq_row)
          
          
          idx = 1:length(f_btrue)
          # Append rows to betaframe
          beta_rows = tibble(
            experiment = i,
            Mc_rep = m,
            beta_id = idx,
            true_beta = f_btrue,
            betahat = f_bhats,
            beta_group = f_gtrue,
            method = "frequentist",
            n_obs = experiment_grid[i,]$n_obs,
            p = experiment_grid[i,]$p,
            g = experiment_grid[i,]$g,
            non_zero = experiment_grid[i,]$non_zero
          )
          all_betas = bind_rows(all_betas, beta_rows)
        }
        
        # Append the true linear predictors
        if (no_error_freq || no_error_brms) {
          idx = 1:length(true_linpreds)
          true_row = tibble(
            experiment = i,
            Mc_rep = m,
            id = idx,
            linpreds = true_linpreds,
            # true_linpreds = f_t_linpreds,
            method = "true",
            n_obs = experiment_grid[i,]$n_obs,
            p = experiment_grid[i,]$p,
            g = experiment_grid[i,]$g,
            non_zero = experiment_grid[i,]$non_zero,
            # True shape and scale used in experiments
            shape = 15,
            scale = 80
          )
          all_experiments = bind_rows(all_experiments, true_row)
        }
        
        # # Get true linpreds from working method
        # true_linpreds = if (no_error_brms) b_t_linpreds else f_t_linpreds
        # 
        # # Create data frame rows
        # new_rows = tibble(experiment = i, Mc_rep = m,
        #                   freq_hat_linpreds = f_h_linpreds,
        #                   bayes_hat_linpreds = b_h_linpreds,
        #                   true_linpreds = true_linpreds,
        #                   bmethod = experiment_grid[i,]$beta_prior,
        #                   n_obs = experiment_grid[i,]$n_obs,
        #                   p = experiment_grid[i,]$p,
        #                   g = experiment_grid[i,]$g,
        #                   non_zero =experiment_grid[i,]$non_zero)
        # 
        
        # Append to all experiment dataframe
    
        # all_experiments = all_experiments %>% bind_rows(new_rows)
      }
    }
    
    #Save output to avoid re-computing
    saveRDS(all_experiments, file = "All_experiments.rds")
    saveRDS(all_betas, file = "All_betas.rds")
  }
  )
all_experiments = readRDS("All_experiments.rds")
all_betas = readRDS("All_betas.rds")
```


```{r} 
# Keep unique parameter cominations
all_experiments = all_experiments %>%
    distinct(Mc_rep, id, method, n_obs, p, g, non_zero, .keep_all = TRUE)
all_betas = all_betas %>%
  distinct(experiment, Mc_rep, beta_id,  method, n_obs, p, g, non_zero, .keep_all = TRUE)
```

```{r}
all_experiments %>%
  filter(g == 4, non_zero == 0.25) %>%
  ggplot(aes(x = method, y = linpreds, fill = method, color = method)) +
  geom_violin(alpha = 0.7, draw_quantiles = c(0.25, 0.5, 0.75)) +
  # geom_jitter(width = 0.1, alpha = 0.5, size = geom_boxplot() +
  geom_vline(xintercept = 3.5, alpha = 1) +
  facet_grid(p ~ n_obs, labeller = labeller(
    p = function(x) paste("p =", x),
    n_obs = function(x) paste("n =", x)
  )) +
  labs(y = "Linear Predictor", x = "Method",
      title = expression("Distribution of the linear predictor "
                        * hat(beta)^T * italic(X) * " compared to true effect "
                        * beta^T * italic(X)),
      subtitle = expression("For g = 4, " * g[nz] * " = 0.25")
      ) +
  theme_linedraw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
all_experiments %>%
  filter(g == 10, non_zero == 0.25) %>%
  ggplot(aes(x = method, y = linpreds, fill = method)) +
  geom_violin(alpha = 0.7, draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_vline(xintercept = 3.5, alpha = 1) +
  facet_grid(p ~ n_obs, labeller = labeller(
    p = function(x) paste("p =", x),
    n_obs = function(x) paste("n =", x)
  )) +
  labs(y = "Linear Predictor", x = "Method",
      title = expression("Distribution of the linear predictor "
                         * hat(beta)^T * italic(X) * " compared to true effect "
                         * beta^T * italic(X)),
      subtitle = expression("For g = 10, " * g[nz] * " = 0.25")
       ) +
  theme_linedraw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
all_experiments %>%
  filter(g == 4, non_zero == 1) %>%
  ggplot(aes(x = method, y = linpreds, fill = method)) +
  geom_violin(alpha = 0.7, draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_vline(xintercept = 3.5) +
  facet_grid(p ~ n_obs, labeller = labeller(
    p = function(x) paste("p =", x),
    n_obs = function(x) paste("n =", x)
  )) +
  labs(y = "Linear Predictor", x = "Method",
      title = expression("Distribution of the linear predictor "
                         * hat(beta)^T * italic(X) * " compared to true effect "
                         * beta^T * italic(X)),
      subtitle = expression("For g = 4, " * g[nz] * " = 1")
       ) +
  theme_linedraw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
all_experiments %>%
  filter(g == 10, non_zero == 1) %>%
  ggplot(aes(x = method, y = linpreds, fill = method)) +
  geom_violin(alpha = 0.7, draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_vline(xintercept = 3.5) +
  facet_grid(p ~ n_obs, labeller = labeller(
    p = function(x) paste("p =", x),
    n_obs = function(x) paste("n =", x)
  )) +
  labs(y = "Linear Predictor", x = "Method",
      title = expression("Distribution of the linear predictor "
                         * hat(beta)^T * italic(X) * " compared to true effect "
                         * beta^T * italic(X)),
      subtitle = expression("For g = 10, " * g[nz] * " = 1")
       ) +
  theme_linedraw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### Shape and scale quality

```{r}

# Remove true (not needed, shape = 15, scale = 80)
est_experiments = all_experiments %>% filter(method != "true") %>%
  distinct(Mc_rep, method, n_obs, p, g, non_zero, .keep_all = TRUE) %>%
  dplyr::select(-c(experiment, id, linpreds)) %>%
  arrange(Mc_rep, n_obs, p, g, non_zero) %>%
  mutate(pg = interaction(p, g, sep = " / "))


# Sparse case
# Top plot (shape)
p1 = est_experiments %>%
  filter(non_zero == 0.25) %>%
  ggplot(aes(x = as.factor(n_obs), y = shape, fill = as.factor(pg))) +
  geom_hline(yintercept = 15, colour = "black", linetype = "dashed") +
  geom_violin() +
  scale_fill_brewer(palette="Dark2") +
  #geom_jitter(alpha = 0.4) +
  facet_wrap(~method) +
  theme_linedraw() +
  theme(legend.position = "none") +
  labs(x = expression(n[obs]), y = "Shape", fill = "p / g") +
  guides(pattern = "none") + labs(title = "Performance of Weibull parameter estimation", 
                                  subtitle = expression("sparse case " * g[nz] == 0.25))

p2 = est_experiments %>%
  filter(non_zero == 0.25) %>%
  ggplot(aes(x = as.factor(n_obs), y = scale, fill = as.factor(pg))) +
  geom_hline(yintercept = 80, colour = "black", linetype = "dashed") +
  geom_violin() +
  scale_fill_brewer(palette="Dark2") +
  #geom_jitter(alpha = 0.4) +
  facet_wrap(~method) +
  theme_linedraw() +
  theme(strip.text.x = element_blank()) +
  labs(x = expression(n[obs]), y = "Scale", fill = "p / g")
  

(p1/p2)
```

```{r}

# Non-sparse case
# Top plot (shape)
p1 = est_experiments %>%
  filter(non_zero == 1) %>%
  ggplot(aes(x = as.factor(n_obs), y = shape, fill = as.factor(pg))) +
  geom_hline(yintercept = 15, colour = "black", linetype = "dashed") +
  geom_violin() +
  scale_fill_brewer(palette="Dark2") +
  #geom_jitter(alpha = 0.4) +
  facet_wrap(~method) +
  theme_linedraw() +
  theme(legend.position = "none") +
  labs(x = expression(n[obs]), y = "Shape", fill = "p / g") +
  guides(pattern = "none") + labs(title = "Performance of Weibull parameter estimation", 
                                  subtitle = expression("non-sparse case " * g[nz] == 1))

p2 = est_experiments %>%
  filter(non_zero == 1) %>%
  ggplot(aes(x = as.factor(n_obs), y = scale, fill = as.factor(pg))) +
  geom_hline(yintercept = 80, colour = "black", linetype = "dashed") +
  geom_violin() +
  scale_fill_brewer(palette="Dark2") +
  #geom_jitter(alpha = 0.4) +
  facet_wrap(~method) +
  theme_linedraw() +
  theme(strip.text.x = element_blank()) +
  labs(x = expression(n[obs]), y = "Scale", fill = "p / g")
  

(p1/p2)
```
## Betas

```{r}
all_betas %>%
  filter(non_zero == 1) %>%
  mutate(g_gnz = interaction(g, non_zero)) %>%
  ggplot(aes(x = true_beta, y = betahat)) +
  geom_point(aes(color = as.factor(g)), alpha = 0.5) +
  scale_shape_manual(values = c(16, 21)) +
  facet_grid(as.factor(p) ~ method,
           labeller = labeller(
             `as.factor(p)` = function(x) paste("p =", x)
           )) +
  geom_vline(xintercept = 0, colour = "black", linetype = "dashed") +
  geom_hline(yintercept = 0, colour = "black", linetype = "dashed") +
  geom_smooth(aes(group = as.factor(n_obs), linetype = as.factor(n_obs)), 
            method = "lm", se = T, size = 0.7, fullrange = TRUE) +
  geom_abline(intercept = 0, slope = 1, colour = "red", linetype = "dashed") +
  theme_linedraw() +
  labs(
    x = expression(beta),
    y = expression(hat(beta)),
    color = "g",
    linetype = expression(n[obs]),
    title = expression(hat(beta) * " versus actual " * beta * " for all experiments in non-sparse case")
  )
```


```{r}
all_betas %>%
  filter(non_zero != 1) %>%
  mutate(g_gnz = interaction(g, non_zero)) %>%
  ggplot(aes(x = true_beta, y = betahat)) +
  geom_point(aes(color = as.factor(g)), alpha = 0.5) +
  scale_shape_manual(values = c(16, 21)) +
  facet_grid(as.factor(p) ~ method,
           labeller = labeller(
             `as.factor(p)` = function(x) paste("p =", x)
           )) +
  geom_vline(xintercept = 0, colour = "black", linetype = "dashed") +
  geom_hline(yintercept = 0, colour = "black", linetype = "dashed") +
  geom_smooth(aes(group = as.factor(n_obs), linetype = as.factor(n_obs)), 
            method = "lm", se = T, size = 0.7, fullrange = TRUE) +
  geom_abline(intercept = 0, slope = 1, colour = "red", linetype = "dashed") +
  theme_linedraw() +
  labs(
    x = expression(beta),
    y = expression(hat(beta)),
    color = "g",
    linetype = expression(n[obs]),
    title = expression(hat(beta) * " versus actual " * beta * " for all experiments in sparse case")
  )
```

# Simulated data

Will show: Betas, X

```{r}
# Import libraries needed to simulate
source("R/generate_data.R")


# Data generation wrapper from main script
generate_data = function(pop_n, obs_n, p, g, ltname, seed, force_recalc, X_plots, 
                         X_rho, β_rho, non_zero_betas, X_scale, active_hazard, 
                         betasep, bscale, betaplot = F, target_snr = 2, 
                         method = "weibull", a, b) {
  
  print("Generating βs...")
  betas = generate_betas(p = p, g=g, rho = β_rho, rho_between = 0,
                       seed = seed, mu_u = betasep, mu_l = -betasep, 
                       beta_scale = bscale, plot = betaplot, 
                       non_zero_groups = non_zero_betas, 
                       active_hazard = active_hazard,
                       target_snr = target_snr,
                       method = method, # Different scaling for different methods
                       weib_shape = a, weib_scale = b,
                       gomp_a = a, gomp_b = b)
  
  # betas$beta = betas$beta + 0.01
  
  
  print("Generating X...")
  X = generate_X(n = pop_n, p = p, g = g, rho = X_rho, seed = seed,
                 rho_between = 0, X_plots = X_plots, scale = X_scale)
  print("Done!")

  
  if (method == "weibull") {
    # a = shape
    # b = scale
    print("Generating Weibull population lifetable...")
    lt = generate_population_lifetable_weibull(N_pop = pop_n, M = p, 
                                              betas = betas$beta, X = X$X, 
                                              filename = ltname, seed = seed, 
                                              force_recalc = force_recalc, 
                                              a = a, b = b)  # Weibull params
    print("Done!")
    
    plot = ggplot(lt, aes(x = t, y = mrl)) +
      geom_line() +
      labs(title = "Weibull Population Lifetable MRL",
           x = "Chronological age (years)", y = "Mean Residual Life") +
      theme_minimal()
    print(plot)
    
    print("Generating data from Weibull lifetable...")
    df_sim = create_dataset_weibull(M = p, n_obs = obs_n, G = g,
                                   gsize = (p/g), lt = lt, betas = betas$beta,
                                   seednr = seed+1, X_plots = F, 
                                   a = a, b = b,  # Weibull params
                                   X_rho = X_rho, X_scale = X_scale, 
                                   followup = 20) # Nr years followed
    
  } else {
    print("Generating Gompertz population lifetable...")
    stop() # Not yet implemented
    lt = generate_population_lifetable_gompertz(N_pop = pop_n, M = p, 
                                               betas = betas$beta, X = X$X, 
                                               filename = ltname, seed = seed, 
                                               force_recalc = force_recalc, 
                                               a = a, b = b)  # Gompertz params
    print("Done!")
    
    plot = ggplot(lt, aes(x = t, y = mrl)) +
      geom_line() +
      labs(title = "Gompertz Population Lifetable MRL",
           x = "Chronological age (years)", y = "Mean Residual Life") +
      theme_minimal()
    print(plot)
    
    print("Generating data from Gompertz lifetable...")
    df_sim = create_dataset_gompretz(M = p, n_obs = obs_n, G = g,
                                    gsize = (p/g), lt = lt, betas = betas$beta,
                                    seednr = seed+1, X_plots = F, 
                                    a = a, b = b,  # Gompertz params
                                    X_rho = X_rho, X_scale = X_scale, 
                                    followup = 20) # Nr years followed
  }
  
  print("Done!")
  return(list(df = df_sim, true_betas = betas, lt = lt))
}


```



```{r}

# p = 50, g = 10, g_nz = 0.5
test_case0.1 = generate_data(pop_n = 10000, obs_n = 1000, p = 50,g = 10,
                           ltname = "plots", seed = 123, force_recalc = T,
                           X_plots = T, X_rho = 0.4, β_rho = 0.4, 
                           non_zero_betas = 0.5, X_scale = 1, active_hazard = 0,
                           betasep = 4, bscale = 1, betaplot = T,
                           target_snr = 2, method = "weibull", a = 15, b = 80)

# p = 500, g = 10, g_nz = 0.5
test_case0.2 = generate_data(pop_n = 10000, obs_n = 1000, p = 500,g = 10,
                           ltname = "plots", seed = 123, force_recalc = T,
                           X_plots = T, X_rho = 0.4, β_rho = 0.4, 
                           non_zero_betas = 0.5, X_scale = 1, active_hazard = 0,
                           betasep = 4, bscale = 1, betaplot = T,
                           target_snr = 2, method = "weibull", a = 15, b = 80)


# p = 50, g = 10, g_nz = 1
test_case1.1 = generate_data(pop_n = 10000, obs_n = 1000, p = 50,g = 10,
                           ltname = "plots", seed = 123, force_recalc = T,
                           X_plots = T, X_rho = 0.4, β_rho = 0.4, 
                           non_zero_betas = 1, X_scale = 1, active_hazard = 0,
                           betasep = 4, bscale = 1, betaplot = T,
                           target_snr = 2, method = "weibull", a = 15, b = 80)

# p = 500, g = 10, g_nz = 1
test_case1.2 = generate_data(pop_n = 10000, obs_n = 1000, p = 500,g = 10,
                           ltname = "plots", seed = 123, force_recalc = T,
                           X_plots = T, X_rho = 0.4, β_rho = 0.4, 
                           non_zero_betas = 1, X_scale = 1, active_hazard = 0,
                           betasep = 4, bscale = 1, betaplot = T,
                           target_snr = 2, method = "weibull", a = 15, b = 80)

# p=g 
# Error in eigen(Sigma, symmetric = TRUE) :
# infinite or missing values in 'x',
# look into this
# So, p -> g

# special_case = generate_data(pop_n = 10000, obs_n = 1000, p = 40, g = 20,
#                            ltname = "plots", seed = 123, force_recalc = T,
#                            X_plots = T, X_rho = 0.4, β_rho = 0.4, 
#                            non_zero_betas = 1, X_scale = 1, active_hazard = 0,
#                            betasep = 0.1, bscale = 1, betaplot = T,
#                            target_snr = 4, method = "weibull", a = 15, b = 80)

```
## Case 0.1


### Beta
```{r}
# p = 50, g = 10, g_nz = 0.5
decimals = 2
b_data = test_case0.1
p = 50
g_nz = 0.5


group_labels = b_data$true_betas %>%
  group_by(group) %>%
  summarise(
    mean_beta = round(mean(beta), decimals),
    sd_beta = round(sd(beta), decimals),
    .groups = 'drop'
  ) %>%
  mutate(label = glue("{group} (μ={mean_beta}, σ={sd_beta})"))

b_data$true_betas %>%
  ggplot(aes(x = as.factor(group), y = beta, fill = as.factor(group),
             colour = as.factor(group))) +
  geom_boxplot(outliers = F, colour = "black") +
  geom_hline(yintercept = 0, linewidth = 0.2) +
  geom_jitter(shape = 21, colour = "black", width = 0.2, size = 1.5) +
  scale_fill_discrete(labels = group_labels$label) +
  theme_linedraw() +
  labs(title = "Distribution of true β-values",
       subtitle = bquote(
         g == 10 ~ " " ~ g[nz] == .(g_nz) ~ " " ~ p == .(p) ~ " | " ~
         mu == .(round(mean(b_data$true_betas$beta), decimals)) ~ " " ~
         sigma == .(round(sd(b_data$true_betas$beta), decimals))
       ),
       x = "Group",
       y = expression(beta),
       fill = "Group Number") +
  guides(colour = "none")
```

### X

```{r}

# Extract covariates (not survival values)
X = as.matrix(test_case0.1$df[, 1:50])

# Overall summary stats
mu = round(mean(X), 2)
sigma = round(sd(X), 2)
corr_mat = cor(X)
corr_long = as.data.frame(as.table(corr_mat))

# Correlation plot of X
p1 = ggplot(corr_long, aes(x = Var1, y = Var2, fill = Freq)) +
  geom_tile() +
  scale_fill_viridis_c() +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
  labs(title = "Correlation and raw values of X",
       subtitle = glue("Overall μ={mu}, σ={sigma}"),
       fill = "Corr")

heatmap_data = as.data.frame(as.table(X))

# X values as-is
p2 = ggplot(heatmap_data, aes(x = Var2, y = Var1, fill = Freq)) +
  geom_tile() +
  scale_fill_viridis_c() +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank()) +
  labs(fill = "Value", y = "Individual", x = "Covariates")

p1 + p2
```


### T_e

```{r}
surv_obj = Surv(time = test_case0.1$df$age_start,
                time2 = test_case0.1$df$age_end,
                event = test_case0.1$df$status)

fit = survfit(surv_obj ~ 1, data = test_case0.1$df)

ggsurvplot(fit,
           title = "Survival of simulated individuals",
           xlab = "Age (years)",
           ylab = "Survival Probability",)

```

## Case 0.2


### Beta
```{r}
# p = 500, g = 10, g_nz = 0.5
decimals = 3
b_data = test_case0.2
p = 500
g_nz = 0.5

group_labels = b_data$true_betas %>%
  group_by(group) %>%
  summarise(
    mean_beta = round(mean(beta), decimals),
    sd_beta = round(sd(beta), decimals),
    .groups = 'drop'
  ) %>%
  mutate(label = glue("{group} (μ={mean_beta}, σ={sd_beta})"))

b_data$true_betas %>%
  ggplot(aes(x = as.factor(group), y = beta, fill = as.factor(group),
             colour = as.factor(group))) +
  geom_boxplot(outliers = F, colour = "black") +
  geom_hline(yintercept = 0, linewidth = 0.2) +
  geom_jitter(shape = 21, colour = "black", width = 0.2, size = 1.5) +
  scale_fill_discrete(labels = group_labels$label) +
  theme_linedraw() +
  labs(title = "Distribution of true β-values",
       subtitle = bquote(
         g == 10 ~ " " ~ g[nz] == .(g_nz) ~ " " ~ p == .(p) ~ " | " ~
         mu == .(round(mean(b_data$true_betas$beta), decimals)) ~ " " ~
         sigma == .(round(sd(b_data$true_betas$beta), decimals))
       ),
       x = "Group",
       y = expression(beta),
       fill = "Group Number") +
  guides(colour = "none")
```



## Case 1.1


### Beta
```{r}
# p = 50, g = 10, g_nz = 1
decimals = 3
b_data = test_case1.1
p = 50
g_nz = 1


group_labels = b_data$true_betas %>%
  group_by(group) %>%
  summarise(
    mean_beta = round(mean(beta), decimals),
    sd_beta = round(sd(beta), decimals),
    .groups = 'drop'
  ) %>%
  mutate(label = glue("{group} (μ={mean_beta}, σ={sd_beta})"))

b_data$true_betas %>%
  ggplot(aes(x = as.factor(group), y = beta, fill = as.factor(group),
             colour = as.factor(group))) +
  geom_boxplot(outliers = F, colour = "black") +
  geom_hline(yintercept = 0, linewidth = 0.2) +
  geom_jitter(shape = 21, colour = "black", width = 0.2, size = 1.5) +
  scale_fill_discrete(labels = group_labels$label) +
  theme_linedraw() +
  labs(title = "Distribution of true β-values",
       subtitle = bquote(
         g == 10 ~ " " ~ g[nz] == .(g_nz) ~ " " ~ p == .(p) ~ " | " ~
         mu == .(round(mean(b_data$true_betas$beta), decimals)) ~ " " ~
         sigma == .(round(sd(b_data$true_betas$beta), decimals))
       ),
       x = "Group",
       y = expression(beta),
       fill = "Group Number") +
  guides(colour = "none")
```


## Case 1.2


### Beta
```{r}
# p = 500, g = 10, g_nz = 1
decimals = 3
b_data = test_case1.2
p = 500
g_nz = 1


group_labels = b_data$true_betas %>%
  group_by(group) %>%
  summarise(
    mean_beta = round(mean(beta), decimals),
    sd_beta = round(sd(beta), decimals),
    .groups = 'drop'
  ) %>%
  mutate(label = glue("{group} (μ={mean_beta}, σ={sd_beta})"))

b_data$true_betas %>%
  ggplot(aes(x = as.factor(group), y = beta, fill = as.factor(group),
             colour = as.factor(group))) +
  geom_boxplot(outliers = F, colour = "black") +
  geom_hline(yintercept = 0, linewidth = 0.2) +
  geom_jitter(shape = 21, colour = "black", width = 0.2, size = 1.5) +
  scale_fill_discrete(labels = group_labels$label) +
  theme_linedraw() +
  labs(title = "Distribution of true β-values",
       subtitle = bquote(
         g == 10 ~ " " ~ g[nz] == .(g_nz) ~ " " ~ p == .(p) ~ " | " ~
         mu == .(round(mean(b_data$true_betas$beta), decimals)) ~ " " ~
         sigma == .(round(sd(b_data$true_betas$beta), decimals))
       ),
       x = "Group",
       y = expression(beta),
       fill = "Group Number") +
  guides(colour = "none")
```


## Lifetables and SNR (p on mrl)

Will do 8 cases:
- snr = 1, p = 200
- snr = 1, p = 20

- snr = 2; p = 200
- snr = 2; p = 20

- snr = 4, p = 200
- snr = 4, p = 20

- snr = 8; p = 200
- snr = 8; p = 20
Then plot in same figure


```{r}
c1 = generate_data(pop_n = 100000, obs_n = 1000, p = 200, g = 10,
                           ltname = "plots", seed = 123, force_recalc = T,
                           X_plots = F, X_rho = 0.4, β_rho = 0.4, 
                           non_zero_betas = 0.5, X_scale = 1, active_hazard = 0,
                           betasep = 4, bscale = 1, betaplot = F,
                           target_snr = 1, method = "weibull", a = 15, b = 80)$lt

c2 = generate_data(pop_n = 100000, obs_n = 1000, p = 20, g = 10,
                           ltname = "plots", seed = 124, force_recalc = T,
                           X_plots = F, X_rho = 0.4, β_rho = 0.4, 
                           non_zero_betas = 0.5, X_scale = 1, active_hazard = 0,
                           betasep = 4, bscale = 1, betaplot = F,
                           target_snr = 1, method = "weibull", a = 15, b = 80)$lt

c3 = generate_data(pop_n = 100000, obs_n = 1000, p = 200, g = 10,
                           ltname = "plots", seed = 125, force_recalc = T,
                           X_plots = F, X_rho = 0.4, β_rho = 0.4, 
                           non_zero_betas = 0.5, X_scale = 1, active_hazard = 0,
                           betasep = 4, bscale = 1, betaplot = F,
                           target_snr = 2, method = "weibull", a = 15, b = 80)$lt

c4 = generate_data(pop_n = 100000, obs_n = 1000, p = 20, g = 10,
                           ltname = "plots", seed = 126, force_recalc = T,
                           X_plots = F, X_rho = 0.4, β_rho = 0.4, 
                           non_zero_betas = 0.5, X_scale = 1, active_hazard = 0,
                           betasep = 4, bscale = 1, betaplot = F,
                           target_snr = 2, method = "weibull", a = 15, b = 80)$lt

c5 = generate_data(pop_n = 100000, obs_n = 1000, p = 200, g = 10,
                           ltname = "plots", seed = 127, force_recalc = T,
                           X_plots = F, X_rho = 0.4, β_rho = 0.4, 
                           non_zero_betas = 0.5, X_scale = 1, active_hazard = 0,
                           betasep = 4, bscale = 1, betaplot = F,
                           target_snr = 4, method = "weibull", a = 15, b = 80)$lt


c6 = generate_data(pop_n = 100000, obs_n = 1000, p = 20, g = 10,
                           ltname = "plots", seed = 128, force_recalc = T,
                           X_plots = F, X_rho = 0.4, β_rho = 0.4, 
                           non_zero_betas = 0.5, X_scale = 1, active_hazard = 0,
                           betasep = 4, bscale = 1, betaplot = F,
                           target_snr = 4, method = "weibull", a = 15, b = 80)$lt

c7 = generate_data(pop_n = 100000, obs_n = 1000, p = 200, g = 10,
                           ltname = "plots", seed = 129, force_recalc = T,
                           X_plots = F, X_rho = 0.4, β_rho = 0.4, 
                           non_zero_betas = 0.5, X_scale = 1, active_hazard = 0,
                           betasep = 0.1, bscale = 1, betaplot = F,
                           target_snr = 8, method = "weibull", a = 15, b = 80)$lt

c8 = generate_data(pop_n = 100000, obs_n = 1000, p = 20, g = 10,
                           ltname = "plots", seed = 130, force_recalc = T,
                           X_plots = F, X_rho = 0.4, β_rho = 0.4, 
                           non_zero_betas = 0.5, X_scale = 1, active_hazard = 0,
                           betasep = 4, bscale = 1, betaplot = F,
                           target_snr = 8, method = "weibull", a = 15, b = 80)$lt

# Case where tSNR is not used p = 20 and p = 200

c9 = generate_data(pop_n = 100000, obs_n = 1000, p = 200, g = 10,
                           ltname = "plots", seed = 131, force_recalc = T,
                           X_plots = F, X_rho = 0.4, β_rho = 0.4, 
                           non_zero_betas = 0.5, X_scale = 1, active_hazard = 0,
                           betasep = 4, bscale = 1/400, betaplot = F,
                           target_snr = 0, method = "weibull", a = 15, b = 80)$lt

c10 = generate_data(pop_n = 100000, obs_n = 1000, p = 20, g = 10,
                           ltname = "plots", seed = 132, force_recalc = T,
                           X_plots = F, X_rho = 0.4, β_rho = 0.4, 
                           non_zero_betas = 0.5, X_scale = 1, active_hazard = 0,
                           betasep = 4, bscale = 1/40, betaplot = F,
                           target_snr = 0, method = "weibull", a = 15, b = 80)$lt

```


```{r}
ps = c(200,20,200,20,200,20,200,20, 200, 20)
tsnrs = c(1,1,2,2,4,4,8,8, "None", "None")
ggplot() +
  geom_line(aes(x = c1$t, y = c1$mrl, colour = as.factor(tsnrs[1]), linetype = as.factor(ps[1]))) +
  geom_line(aes(x = c2$t, y = c2$mrl, colour = as.factor(tsnrs[2]), linetype = as.factor(ps[2]))) +
  geom_line(aes(x = c3$t, y = c3$mrl, colour = as.factor(tsnrs[3]), linetype = as.factor(ps[3]))) +
  geom_line(aes(x = c4$t, y = c4$mrl, colour = as.factor(tsnrs[4]), linetype = as.factor(ps[4]))) +
  geom_line(aes(x = c5$t, y = c5$mrl, colour = as.factor(tsnrs[5]), linetype = as.factor(ps[5]))) +
  geom_line(aes(x = c6$t, y = c6$mrl, colour = as.factor(tsnrs[6]), linetype = as.factor(ps[6]))) +
  geom_line(aes(x = c7$t, y = c7$mrl, colour = as.factor(tsnrs[7]), linetype = as.factor(ps[7]))) +
  geom_line(aes(x = c8$t, y = c8$mrl, colour = as.factor(tsnrs[8]), linetype = as.factor(ps[8]))) +
  # No SNR scaling
  geom_line(aes(x = c9$t, y = c9$mrl, colour = as.factor(tsnrs[9]), linetype = as.factor(ps[9]))) +
  geom_line(aes(x = c10$t, y = c10$mrl, colour = as.factor(tsnrs[10]), linetype = as.factor(ps[10]))) +
  labs(y = "Mean residual life", x = "Chronological age (years)", colour = expression(t[snr]),
       linetype = "p", title = "Comparison of population lifetables",
       subtitle = expression("For various p and " * t[snr])) +
  theme_linedraw()
```

