---
title: "survRRR external validation using NEO"
author: "Philippe A.N. Berends"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

# Set up project

## Load packages

```{r, results='hide', message=FALSE, warning=FALSE}
# (Down)load necessary packages
# install.packages("parallelly") # otherwise other installs will break
# install.packages("glmnet")
# install.packages("mstate")
# install.packages("tidyverse")
# install.packages("RecordLinkage")
# install.packages("survminer")
# install.packages("ggsurvfit")
# install.packages("future.apply") # Needed to install RecordLinkage

# Load survRRR
library(tidyverse)
library(readxl)   # for supplementary excel files
library(haven) # to read the .dta
library(glue) # variables in strings
library(survminer)
source("penalized_survRRR-main/penalized_survRRR-main/survRRR_functions.R")
library(survival)
library(ggsurvfit)  # From survminer package
source("utils.R") # custom functions for preprocessing
```

## Load data from secure drive

```{r}
# Load dataset
data <- read_dta("//vf-DataSafe/DataSafe$/Div0/ict/survRRR in NEO_2873/NEO_analyse_PPXX_metabolytes_morbidity_RodriquezGirondo_2025-02-18.dta")
# add id for ease of use
data <- data %>% mutate(id = 1:nrow(data))
```


# Pre-processing


## Select intersection of metabolites between NEO and UK biobank

Use Levenshtein similarity to get similarities between pre-processed strings (removed special characters)
Afterwards a manual comparison between non-100% similar metabolites

```{r}
# Select only the metabolite variables
metabolites = data[64:ncol(data)]
# grab labels from STATA file
metabolites_labels = lapply(metabolites, function(x) attributes(x)$label) %>% as.character()
metabolites_labels_unique = metabolites_labels %>%
  # Remove the first word (split on first space and take everything after)
  map_chr(~ sub("^\\S+\\s+", "", .)) %>%
  unique()


# Derive cleaner names from the metabolic columns
cleaned_labels = sapply(metabolites_labels_unique, function(text) {
  # remove everything including and after " - "
  result = sub("\\ - .*$", "", text)
  # remove words with a dot
  result = gsub("\\S*\\.\\S*", "", result)
  2
  # Clean up any trailing whitespace
  result = trimws(result)
  return(result)
}) %>% unique()


pre_meal_metab = metabolites %>% select(matches("_1"))
post_meal_metab = metabolites %>% select(matches("_3"))



reference = "~/survRRR validation/supporting_information/SuppInfo_Table1_metabolic_variables_description.xlsx"
matchframe = exctract_intersect_metabolites(cleaned_labels = cleaned_labels, reference = reference)
```

Now we have a dataframe called "matchframe" we can use to translate NEO metabolites to UKbiobank



## Turn dataframe to long format

### Dataset pre-processing

```{r}
pre_meal_covs = pre_meal_metab[,matchframe$true.match == 1] %>%
  mutate(id = row_number())
p = pre_meal_covs %>% ncol() -1 # minus the id column

p # ncol (metabolites)

# number of people in dataset pre-preprocessing
dim(pre_meal_covs)


# Keep patients above age 50, to quote the pen survRRR paper:
# "Furthermore, participants 305
# were only included if they were 50 years or older at time of recruitment, as those below 50 years of age 306
# were considered to be too young to be diagnosed with an age-related disease during the follow-up period."
pre_meal_covs = pre_meal_covs[data$leeftijd >= 50,]
dim(pre_meal_covs)


# # ------ handle missing data --- (all metabolites have >0.01 missing measurements)
# 
# # Remove people with more than 1% missing metabolite measurements
# # NOTE: Step NOT performed in original paper, but if I don't do this, ALL bio markers
# # will have more than 1% missing measurements. (e.g., df will be empty in step hereafter)
# pre_meal_covs <- pre_meal_covs[!(((rowSums(pre_meal_covs %>% is.na()))/p) > 0.06),]
# 
# #nrow(pre_meal_covs)
# 
# 
# # Remove metabolites with more than 1% missing (using pre_meal_covs data)
# pre_meal_covs <- pre_meal_covs[,!(((colSums(pre_meal_covs %>% is.na()))/nrow(pre_meal_covs)) > 0.01)]
# 
# ncol(pre_meal_covs) - 1
# 
# # Remove people with missing metabolite values
# pre_meal_covs <- pre_meal_covs[!(((rowSums(pre_meal_covs %>% is.na()))/p) > 0.00),]
# nrow(pre_meal_covs)
# 

# NEW: Replace missing metabs with min(metab column)/2
# - apply column where is na = min column / 2

# Get the min/2 value for each column separately
pmc_min2 = pre_meal_covs %>% apply(MARGIN = 2, FUN = function(x) min(x, na.rm = T))/2

# Insert these into the missing values on column-level
pre_meal_covs = pre_meal_covs %>% replace_na(as.list(pmc_min2))

# post_meal_covs = post_meal_metab[,matchframe$true.match == 1] %>%
#   mutate(id = row_number())
# data parameters
# outcomes: CVA, CVD, diabetes, MI, Stroke, TIA, (DEATH)

# Hypertension can not be included due to lack of followup data
# Intersect between original paper and these data:
#       TIA, MI, DB, Death
k = 4


# outcome dataframe for diabetes
outcome_diabetes <- data %>%
  select(id, visitdd, diab_prev, diabetes2, diabetes2_date, einddatum2) %>%
  mutate(outcome_DIA_date = ifelse(diab_prev == 0 & diabetes2 == 1, diabetes2_date, 0),
         outcome_DIA = ifelse(outcome_DIA_date > 0, 1, 0),
         censored_DIA = ifelse(outcome_DIA == 1, 0, 1)
         )

# Outcome dataframe for MI
outcome_MI <- data %>%
  select(id, visitdd, MI2_Date_1, MI2_Date_2, MI2_Date_3, MI2_inc, einddatum2)
# Outcome dataframe for Stroke
outcome_stroke <- data %>%
  select(id, visitdd, Stroke2_date_1, Stroke2_date_2, Stroke2_date_3, Stroke2_inc, einddatum2) %>%
  mutate(
    valid_dates = pmin(Stroke2_date_1, Stroke2_date_2, Stroke2_date_3, na.rm = TRUE),
    outcome_STR = ifelse(Stroke2_inc == 1 | (!is.na(valid_dates) & valid_dates > visitdd), 1, 0),
    outcome_STR_time = ifelse(!is.na(valid_dates) & valid_dates > visitdd, valid_dates, NA),
    censored_STR = ifelse(is.na(Stroke2_date_1) & is.na(Stroke2_date_2) & is.na(Stroke2_date_3), 
                      1, 
                      ifelse(!is.na(valid_dates) & valid_dates <= visitdd, 1, 0))
  ) %>%
  select(-c(valid_dates))

# Outcome dataframe for TIA
outcome_TIA <- data %>%
  select(id, visitdd, TIA2_date_1, TIA2_date_2, TIA2_date_3, TIA2_inc, einddatum2)

# ?outcome dataframe for death? 
outcome_death <- data %>%
  select(id, visitdd, einddatum2, eind2) %>% #4 = death
  mutate(outcome_DTH = ifelse(eind2 == 4, 1, 0),
         outcome_DTH_date = ifelse(outcome_DTH == 1, einddatum2, 0),
         censored_DTH = ifelse(outcome_DTH == 1, 0, 1))
# ts and the ds (censored)


# data frame with relevant covariates
# covariates <- data %>%
#   select(id, leeftijd, sexe, bmim, bpsystgem, bpdiastgem)


# # exclude people who have NA for final date (still in study)
# outcome_dataframe = outcome_dataframe %>% subset(!is.na(einddatum2))
# Create empty dataframe to store results
```


```{r}
head(outcome_death)
```

```{r}
head(outcome_MI)
```

```{r}
head(outcome_TIA)
```

DIA outcome is directly taken from the data.


### Create long format dataframe

```{r}
source("utils.R") # custom functions for preprocessing

# For UKB likeness
pre_dlong = make_long(occasion = 1, data = data, outcome_death = outcome_death, outcome_MI = outcome_MI, outcome_TIA = outcome_TIA, meal_covs = pre_meal_covs, standardize = T, filename = "pre_dlong_standardized_ratios-standardised", log_transform = F)


#pre_dlong = make_long(occasion = 1, data = data, outcome_death = outcome_death, outcome_MI = outcome_MI, outcome_TIA = outcome_TIA, meal_covs = pre_meal_covs, standardize = F, filename = "pre_dlong_unstandardized")

# Data consists of only included individuals in dlong
data_filt = data %>% filter(id %in% (pre_dlong$id %>% unique()))
# Remove no longer relevant variables to save space
# rm(pre_meal_metab)
# rm(outcome_death)
# rm(outcome_diabetes)
# rm(outcome_MI)
# rm(outcome_stroke)
# rm(outcome_TIA)

head(pre_dlong[,1:7])
```

## Exploratory data analysis

Various plots that give summary statistics of the individuals from the original dataframe that have been included in the long data.

Distribution of age/gender with years in study

```{r}
data_filt %>% mutate(sex = case_when(
  sexe == 1 ~ as.factor("Male"),
  sexe == 2 ~ as.factor("Female"))
  ) %>% 
  pull(sex) %>%
  table() %>%
  addmargins(1)

data_filt %>% mutate(sex = case_when(
  sexe == 1 ~ as.factor("Male"),
  sexe == 2 ~ as.factor("Female"))
  ) %>%
  ggplot(aes(x = leeftijd, fill = sex)) +
  geom_bar(position = "dodge") +
  labs(x = "Age", y = "Count", title = "Age at inclusion Distribution by Sex") +
  theme_minimal()

data_filt %>% mutate(sex = case_when(
  sexe == 1 ~ as.factor("Male"),
  sexe == 2 ~ as.factor("Female"))
  ) %>%
  ggplot(aes(x = bmim, fill = sex)) +
  geom_histogram(position = "dodge", binwidth = 1) +
  labs(x = "BMI", y = "Count", title = "Distribution of BMI by Sex") +
  theme_minimal()


```

```{r}
# DIA, DTH, MI, TIA
pre_dlong %>% 
    mutate(outcome = case_when(
        trans == 1 ~ "DIA",
        trans == 2 ~ "Death",
        trans == 3 ~ "MI",
        trans == 4 ~ "TIA",
        trans == 5 ~ "FIRST"
    )) %>% select(c(outcome, status)) %>% table %>%
    addmargins(c(1,2))

# Create the plot with Kaplan-Meier curves for each outcome

survfit = survfit2(Surv(Tstart, Tstop, status) ~ outcome, data = pre_dlong %>% 
    mutate(outcome = case_when(
        trans == 1 ~ "DIA",
        trans == 2 ~ "Death",
        trans == 3 ~ "MI",
        trans == 4 ~ "TIA",
        trans == 5 ~ "FIRST"
    )))
p = ggsurvfit(survfit) +
  xlab("Age")
p
```


What was the first event on an individual-level?

```{r}
# Look at trans 1, 2, 3, 4
tabs = pre_dlong %>% filter(status == 1, trans != 5) %>% # The event should have occurred
  group_by(id) %>% # Look at individual-level
  slice_min(Tstop, n = 1) %>% # Take the first event based on Tstop (time since start)
  select(trans) %>% # Look at events 1, 2 3, 4 ("DIA", "Death", "MI", "TIA")
  table %>%
  colSums() %>% # Count how often the event in question was the first (each event a column)
  as.data.frame() %>% t # Transpose because R likes to make rows

colnames(tabs) = c("DIA", "Death", "MI", "TIA")
rownames(tabs) = "first_event"
tabs
```

# External validation

## Import UKbiobank data

```{r}
# import alphas and betas from UKBiobank model
uk_alpha = read_excel("~/survRRR validation/supporting_information/SuppInfo_Table6_Alpha_hat.xlsx")
uk_beta = read_excel("~/survRRR validation/supporting_information/SuppInfo_Table5_Beta_hat.xlsx")

# remove _. from name
pre_dlong <- pre_dlong %>%
  rename_with(~ gsub("_\\d+$", "", .), matches("_\\d+$"))

# small dataframe to remap names with
name_mapping <- setNames(
  matchframe$Metabolic.variable,  # replace by
  matchframe$column_name # from dlong
)

columns_to_rename <- intersect(names(pre_dlong), names(name_mapping))
pre_dlong <- pre_dlong %>%
  rename_at(vars(columns_to_rename), ~ name_mapping[.])

names(pre_dlong[,8:ncol(pre_dlong)])  %>% length
uk_alpha$metabolic_variable %>% length

intersect(names(pre_dlong), uk_alpha$metabolic_variable) %>% length

# these metabolites don't occur in the alphas
setdiff(names(pre_dlong), uk_alpha$metabolic_variable) # 7 don't occur, as survRRR was not fit on these! I checked and the paper confirms it

# thus we exclude them from the long frame
pre_dlong <- pre_dlong %>% select(-setdiff(names(pre_dlong), uk_alpha$metabolic_variable)[9:length(setdiff(names(pre_dlong), uk_alpha$metabolic_variable))])

setdiff(names(pre_dlong),c("status", "Tstart", "Tstop", "trans", 
                             "from", "to", "id", "X")) %>% length()

```

## Calculate C-index

```{r}
usable_uk_beta = uk_beta[uk_beta$metabolic_variable %in% names(pre_dlong),]
usable_uk_alpha = uk_alpha[uk_alpha$metabolic_variable %in% names(pre_dlong),]
outcome_map = c("diabetes", "death", "mi", "tia", "first")



# Calculate C-index for each outcome
c_indexes = list()
for (i in 1:length(outcome_map)) {
  c_indexes[[outcome_map[i]]] = calculate_cindex(pre_dlong, i, usable_uk_alpha, outcome_map)
}
```

 ## Compare with MetaboHealth score

```{r}
# For MetaboHealth (log transformed)
pre_dlong_MH = make_long(occasion = 1, data = data, outcome_death = outcome_death, outcome_MI = outcome_MI, outcome_TIA = outcome_TIA, meal_covs = pre_meal_covs, standardize = T, filename = "pre_dlong_standardized_log_ratios-standardised", log_transform = T)

# remove _. from name
pre_dlong_MH <- pre_dlong_MH %>%
  rename_with(~ gsub("_\\d+$", "", .), matches("_\\d+$"))


# select the correct biomarker columns
# XXL-VLDL-L, S-HDL-L, VLDL-D, PUFA/FA, Glc, Lac, His, Ile, Leu, Val, Phe, AcAce, ALb, GlycA

# get hazard ratios from table at: https://www.nature.com/articles/s41467-019-11311-9/tables/2
metabohealth = data.frame(metabolic_variable = c('XXL-VLDL-L', 'S-HDL-L', 'VLDL-D', 'PUFA/FA', 'Glc', 'Lac', 'His', 'Ile', 'Leu', 'Val', 'Phe', 'AcAce', 'ALb', 'GlycA'), coefficient = c(0.80, 0.87, 0.85, 0.78, 1.16, 1.06, 0.93, 1.23, 0.82, 0.87, 1.13, 1.08, 0.89, 1.32))

# transform to log hazards
metabohealth$coefficient = log(metabohealth$coefficient)

# Select right biomarkers from long frame
metabohealth_biomarkers = pre_dlong_MH %>% select(c(id, Tstart, Tstop, status, from, to, trans, XXLVLDLL, SHDLL,
                                                 VLDLD, PUFAFA, Glc, Lac, His,
                                                 Ile, Leu, Val, Phe,
                                                 Alb, Gp)) # Acetoacetate not available in NEO

# remove AcAce, not available in NEO
metabohealth = metabohealth[-12,]

# set names equal to MetaboHealth
names(metabohealth_biomarkers)[8:ncol(metabohealth_biomarkers)] = metabohealth$metabolic_variable

# Calculate C-index
c_indexes_MH = list()
for (i in 1:length(outcome_map)) {
  c_indexes_MH[[outcome_map[i]]] = calculate_cindex(metabohealth_biomarkers, i, metabohealth, outcome_map)
}

c_indexes_MH

```


# Correlation linpred survRRR with Metabohealth


```{r}
# unique covariates and individuals
corrset_S = pre_dlong %>% distinct(id, .keep_all = T)
corrset_M = metabohealth_biomarkers %>% distinct(id, .keep_all = T)

lps_survRRR = as.vector(as.matrix(
  corrset_S[, intersect(
    colnames(corrset_S), usable_uk_alpha$metabolic_variable)
    ]) %*% usable_uk_alpha$coefficient)

lps_MetaboHealth = as.vector(as.matrix(
  corrset_M[, intersect(
    colnames(corrset_M), metabohealth$metabolic_variable)
    ]) %*% metabohealth$coefficient)

library(ggpubr)
corrset_All = data.frame(MetaboHealth = lps_MetaboHealth, survRRR = lps_survRRR)
corrset_All %>% ggplot(aes(x = MetaboHealth, y = survRRR)) +
  geom_point() +
  geom_abline(, slope = 1) +
  geom_smooth(method = lm)+
  stat_cor(method = "pearson") +
  theme_minimal() +
  labs(title = "Correlation between MetaboHealth and survRRR linear predictors")
```